---
title: "ESM 244 Lab 8"
author: "Giovanna Davila"
date: "3/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Load the packages

```{r packages}

library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(spatstat)
library(maptools)

```

###Column graph of Texas oil spills

```{r load_data}

oil_spills <- read_csv("oil_spills.csv")

df <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` < 2017) %>% 
  group_by(`Accident Year`) %>% 
  summarise(Loss = sum(`Net Loss (Barrels)`))

colnames(df) <- c("Year","Loss")

ggplot(df, aes(x = Year, y = Loss)) +
  geom_col()

```


###Leaflet plot of spill locations in Texas in 2016

```{r}

df_loc <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` == 2016) %>% 
  select(Latitude, Longitude, `Net Loss (Barrels)`)

colnames(df_loc) <- c("latitude","longitude","net_loss")

oil_sf <- st_as_sf(df_loc, coords = c("longitude","latitude"), crs = 4326) # Leaflet does recognize longitude/latitude without having to convert to spatial features, leaflet will just guess what your longitude latitude is when not in spatial features

leaflet(oil_sf) %>% 
  addTiles() %>%
  addMarkers()

```

### tmap plot with the Texas state shapefile

```{r}

states <- st_read(dsn = ".", layer = "states")

tex_border <- states %>% 
  filter(STATE_NAME == "Texas") %>% 
  st_transform(4326)

plot(tex_border)

tm_shape(tex_border) +
  tm_polygons() +
  tm_shape(oil_sf) +
  tm_dots(size = 0.3)

```

Question: 
Are the points spatially random? 


###Convert the data to spatial points patterns (combination of the point data and the bounding window)

```{r}

# Convert back from simple features to a data frame, but still not recognized as something to do point pattern analysis with

spill_sp <- as(oil_sf, "Spatial")

spill_ppp <- as(spill_sp, "ppp")

tx_sp <- as(tex_border, "Spatial")
tx_owin <- as(tx_sp, "owin")
# Setting this as the outer window for the spatial point analysis


all_ppp <- ppp(spill_ppp$x, spill_ppp$y, window = tx_owin)
# Some points ommitted that don't align with the window 

```


###A density plot:

```{r}

# Taking our combined point pattern info and creating a density plot
plot(density(all_ppp, sigma = 0.4))

```

Recall: quadrat test really tests spatial evenness (even though its called a test of spatial randomness)

## Quadrat test for spatial evenness

Are oil spills spatially distrubuted throughout the state?

```{r}

oil_qt <- quadrat.test(all_ppp, nx = 5, ny = 5)

# Separates Texas into quadrats, counts up the observations in each, performs Chi Square
# Testing a null hypothesis of complete spatial randomness

oil_qt

# Retain alernative hypothesis that data are not evenly distributed

plot(all_ppp)
plot(oil_qt, add = TRUE, cex = 0.4)

# Automatically calculates area and proportional number of counts in smaller quadrats based on the shape of Texas
#In the graph, are given expected counts, standard deviations versus expected value and total quadrat area value

```

If data are spatially clustered, then nearest neighbors would be closer than they would be under CSR

###G-Function for nearest neighbor analysis

```{r}

r <- seq(0,1, by = 0.01) # This sequence makes sense for the data, won't always be the case
# Lag distance between points

# Based on poisson distribution, recognizing the window and the points, what would it look like as CSR data?

oil_gfun <- envelope(all_ppp, fun = Gest, r = r, nsim = 100) # Run 100 simulations using the r sequence and the G function

# Theo column = theoretical modeled CSR

ggplot(oil_gfun, aes(x = r, y = obs)) +
  geom_line(color = "black") +
  geom_line(aes(x = r, y = theo), color = "red")

# Our observed data has a higher proportion of point pairs w/ nearest neighbors at shorter distances compared to CSR data

```

###Nearest neighbor using the L-function (Ripley's K, standardized)

```{r}

# This is MORE computationally intensive. Looks at every event (point) in spcae and makes increasing bubbles around it until it incorporates all points around it that exist

r2 <- seq(0,3, by = 0.5)
# Chose a large increment so simulations wouldn't run forever

oil_lfun <- envelope(all_ppp, fun = Lest, r = r2, nsim = 20, global = TRUE)

ggplot(oil_lfun, aes(x = r2, y = obs)) +
  geom_line(color = "green") +
  geom_line(aes(x = r2, y = theo), color = "blue")


```

Conclusions: 
- Map showed the data didn't look CSR
- Ran a test that indicated it was not CSR
- Both the G and L functions indicated the data was not CSR






